Approximation as Differentiation

### 2025-07-10 main/language.tex
WIP.

A model $M$ of $\Sigma$ can be given in any category $\cat{C}$ with finite products and a terminal object, and
assigns to each base type $\rho \in \PrimTy$ an object $\sem{\rho}_{M}$ in $\cat{C}$, and to each primitive
operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\phi}_{M}: \sem{\rho_1}_{M} \times
\ldots \times \sem{\rho_n}_{M} \to \sem{\rho}_{M}$. A model of $\tyList$ can also be given in $\cat{C}$ and
consists of an endofunctor $\List: \cat{C} \to \cat{C}$, plus for any object $X$, morphisms $\nil: 1 \to
\List(X)$ and $\cons: X \times \List(X) \to \List(X)$, and for any objects $Y, Z$ and morphisms $f_\nil: Z \to
Y$ and $f_\cons: Z \times X \times Y \to Y$, a morphism $\fold(f_\nil,f_\cons): Z \times \List(X) \to Y$.

To give the semantics for the full language, we fix a bicartesian closed category $\Sem$ with a strong monad
$(\Lift, \eta, \mu)$ with strength $\mathsf{st}$, a model $M$ of $\Sigma$ in $\Sem$, and a model of $\tyList$
in $\Sem$.

\figref{semantics:types} gives the interpretations of types $\tau$ and contexts $\Gamma$ as objects of $\Sem$.
\figref{semantics:terms} gives the interpretation of terms $\Gamma \vdash t: \tau$ as morphisms $\sem{\Gamma}
\to \sem{\tau}$.

>>>>

A model $M$ of $\Sigma$ can be given in any category $\cat{C}$ with finite products and a terminal object, and
assigns to each base type $\rho \in \PrimTy$ an object $\sem{\rho}_{M}$ in $\cat{C}$, and to each primitive
operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\phi}_{M}: \sem{\rho_1}_{M} \times
\ldots \times \sem{\rho_n}_{M} \to \sem{\rho}_{M}$. A model of $\tyList$ can also be given in such a category,
and consists of an endofunctor $\List: \cat{C} \to \cat{C}$, plus for any object $X$, morphisms $\nil: 1 \to
\List(X)$ and $\cons: X \times \List(X) \to \List(X)$, and for any objects $Y, Z$ and morphisms $f_\nil: Z \to
Y$ and $f_\cons: Z \times X \times Y \to Y$, a morphism $\fold(f_\nil,f_\cons): Z \times \List(X) \to Y$,
satisfying the usual laws for lists.

To give the semantics for the full language, we fix a bicartesian closed category $\Sem$ with a strong monad
$(\Lift, \eta, \mu)$ with strength $\mathsf{st}$, a model $M$ of $\Sigma$ in $\Sem$, and a model of $\tyList$
in $\Sem$. \figref{semantics:types} gives the interpretations of types $\tau$ and contexts $\Gamma$ as objects
of $\Sem$. \figref{semantics:terms} gives the interpretation of terms $\Gamma \vdash t: \tau$ as morphisms
$\sem{\Gamma} \to \sem{\tau}$.

### 2025-07-10 main/models.tex
WIP.

\begin{proposition}
  \label{prop:ho-embedding}
  The functor
  $F : \Fam(\LatGal) \to \Fam(\MeetSLat \times \JoinSLat^\op)$ is defined
  on objects as
  $F(X, \partial X) = (X, \lambda x. (\partial X(x), \partial
  X(x)))$. This functor is faithful and preserves coproducts and
  finite products.
\end{proposition}

>>>>

\begin{proposition}
  \label{prop:ho-embedding}
  The functor
  $\HoEmbed : \Fam(\LatGal) \to \Fam(\MeetSLat \times \JoinSLat^\op)$ is defined
  on objects as
  $\HoEmbed(X, \partial X) = (X, \lambda x. (\partial X(x), \partial
  X(x)))$. This functor is faithful and preserves coproducts and
  finite products.
\end{proposition}
